---
title: "Stack-Tags Analysis"
author: "Kyle MacDonald"
date: "11/25/2016"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=T, warning=F, cache=F, message=F, sanitize = T)
library(lme4)
library(tidyverse)
library(magrittr)
library(knitr)
library(stringr)
library(lubridate)
theme_set(theme_bw())
```

## Load and clean dataset

These data are organized into three tables:

* Questions contains the title, body, creation date, score, and owner ID for each R question.
* Answers contains the body, creation date, score, and owner ID for each of the answers to these questions. The ParentId column links back to the Questions table.
* Tags contains the tags on each question besides the R tag.

```{r load_data, echo = FALSE, cache = T}
#answers <- read_csv("../data/Answers.csv", progress = FALSE)
questions <- read_csv("../data/Questions.csv", progress = FALSE)
tags <- read_csv("../data/Tags.csv", progress = FALSE)
```

Merge question information with tags.

```{r}
tags <- questions %>% 
  select(Id, Score, OwnerUserId, CreationDate) %>% 
  left_join(tags, ., by = "Id")
```

Split date-time variable.

```{r}
tags %<>% 
  mutate(yr = year(CreationDate),
         mon = month(CreationDate), 
         dy = day(CreationDate))
```

Filter dataset to remove questions with extreme scores and to remove the year 2008, which did not have very much data.

```{r}
tags %<>%
   filter(Score <= mean(Score) + 2*sd(Score), 
         Score >= mean(Score) - 2*sd(Score),
         year(CreationDate) > 2008)
```

## Descriptives 

Get the distribution of questions over number of tags.

```{r}
tags %<>% 
  group_by(Id) %>% 
  summarise(n.tag = n()) %>% 
  left_join(tags, ., by = "Id")
```

Now plot the proportion of questions that used 1-4 tags.

```{r}
tags %>%
  group_by(n.tag) %>% 
  summarise(tag.prop = n() / nrow(.)) %>% 
  ggplot(aes(x = n.tag, y = tag.prop)) +
  geom_bar(stat = "identity")
```

Most questions have 1 or 2 tags. 

## Tag analysis

#### Does the number of tags affect the question score?

```{r}
tags %>% 
  ggplot(aes(x = n.tag, y = Score, group = n.tag)) + 
  geom_boxplot()
```

Nope!

#### Has the number of tags people use changed over time?

```{r}
tags %>% 
  group_by(mon, yr) %>% 
  summarise(m.tag = mean(n.tag)) %>% 
  ggplot(aes(x = mon, y = m.tag, color = as.factor(yr), group = yr)) +
  geom_line()
```

No changes over the course of a year, which makes sense. And it looks like 2009 is noisy or when tags were first introduced, meaning the community was still figuring out how to use them. 

What if we collapse across months to just look at the change over years.

```{r}
tags %>% 
  group_by(yr) %>% 
  summarise(m.tag = mean(n.tag),
            n = n()) %>% 
  ggplot(aes(x = yr, y = m.tag)) +
  geom_smooth(se=F) +
  geom_point(aes(size = n)) + 
  ylim(1, 4)
```

Maybe a slight increase in tag use, but not much going on here.

```{r}
m1 <- lmer(n.tag ~ scale(yr) + (1|OwnerUserId), data = tags)

broom::tidy(m1) %>% kable(digits = 3)
```

Yep, very small, but reliable effect of year: tag use is increasing over time, but in very small increments that are likely not practically meaningful.


## Dimensionality reduction on tags

TODO: figure out how to compress the number of tags using some kind of unsupervised learning algorithm, e.g., cluster analysis or PCA. 
