---
title: "Stack-Questions Analysis"
author: "Kyle MacDonald"
date: "11/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=F, cache=F, message=F, sanitize = T)
library(tidyverse)
library(knitr)
theme_set(theme_bw())
```

Load data. This is organized as three tables:

* Questions contains the title, body, creation date, score, and owner ID for each R question.
* Answers contains the body, creation date, score, and owner ID for each of the answers to these questions. The ParentId column links back to the Questions table.
* Tags contains the tags on each question besides the R tag.

```{r load_data, echo = FALSE, cache = T}
#answers <- read_csv("../data/Answers.csv", progress = FALSE)
questions <- read_csv("../data/Questions.csv", progress = FALSE)
#tags <- read_csv("../data/Tags.csv", progress = FALSE)
```

### Descriptives and sanity checks

```{r n_users}
n.users <- questions %>% 
  select(OwnerUserId) %>% 
  unique() %>% 
  count()

n.questions <- nrow(questions)
```

There are `r n.users` unique users and `r n.questions` questions in the dataset.

### Score analysis

What is the distribution over the *Score* variable? 

```{r score_dist}
ggplot(aes(x = Score), data = questions) + 
  geom_histogram()
```

Looks like there are a couple of extreme values here, preventing us from getting a feel for the distribution. We can filter scores that are greater than +/- 2SD from the mean of score.

```{r}
questions %>% 
  filter(Score <= mean(Score) + 2*sd(Score), 
         Score >= mean(Score) - 2*sd(Score)) %>% 
  ggplot(aes(x = Score), data = .) + 
  geom_histogram(bins = 30)
```

This distribution is looking more reasonable. But what does it mean to get negative score? Let's look at a small sample of the questions with the lowest scores to see what's going on. 

```{r sample_neg_questions}
worst.qs <- questions %>% 
  select(Score, Body) %>%
  arrange(Score) %>% 
  head(n=5) 
```

```{r sample_best_questions}
best.qs <- questions %>% 
  select(Score, Body) %>%
  arrange(desc(Score)) %>% 
  head(n=5) 
```

TODO: figure out the best way to render the html in the Body variable. 

### Looking for users

What is the distribution of questions over users? First we compute some basic descriptives.

```{r}
ss.qs <- questions %>% 
  filter(is.na(OwnerUserId) == F) %>% 
  group_by(OwnerUserId) %>% 
  count() 

q.sum <- summary(ss.qs$n)
```

The average number of questions per user is `r summary(ss.qs$n)["Mean"]`. The min is `r summary(ss.qs$n)["Min."]` (makes sense) and the max is `r summary(ss.qs$n)["Max."]` (wow, that's a lot of questions!).

Let's take a look at the person who asked 357 questions to see what kind of questions they are asking and over what time frame they asked these questions. 

```{r}
max.q.id <- ss.qs %>% 
  filter(n == summary(ss.qs$n)["Max."]) %>% 
  select(OwnerUserId) %>% 
  as.character()

ss.max <- questions %>% 
  filter(OwnerUserId == max.q.id)
```


