---
title: "Stack-Answers Analysis"
author: "Kyle MacDonald"
date: "11/25/2016"
output: html_document
---
---
title: "Stack-Tags Analysis"
author: "Kyle MacDonald"
date: "11/25/2016"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=T, warning=F, cache=F, message=F, sanitize = T)
library(lme4)
library(tidyverse)
library(magrittr)
library(knitr)
library(stringr)
library(lubridate)
theme_set(theme_bw())
```

## Load and clean dataset

These data are organized into three tables:

* Questions contains the title, body, creation date, score, and owner ID for each R question.
* Answers contains the body, creation date, score, and owner ID for each of the answers to these questions. The ParentId column links back to the Questions table.
* Tags contains the tags on each question besides the R tag.

```{r load_data, echo = FALSE, cache = T}
answers <- read_csv("../data/Answers.csv", progress = FALSE)
questions <- read_csv("../data/Questions.csv", progress = FALSE)
#tags <- read_csv("../data/Tags.csv", progress = FALSE)
```

Merge question information with tags.

```{r}
answers <- questions %>% 
  select(ParentId = Id, Score.q = Score, 
         OwnerUserId.q = OwnerUserId, 
         CreationDate.q = CreationDate) %>% 
  left_join(answers, ., by = "ParentId")
```

Split date-time variable for both questions and answers.

```{r}
answers %<>% 
  mutate(yr.a = year(CreationDate),
         mon.a = month(CreationDate), 
         dy.a = day(CreationDate),
         yr.q = year(CreationDate.q),
         mon.q = month(CreationDate.q), 
         dy.q = day(CreationDate.q))
```

Filter dataset. Remove questions and answers with extreme scores and remove the year 2008, which did not have very much data.

```{r}
answers %<>%
   filter(Score <= mean(Score) + 2*sd(Score), 
         Score >= mean(Score) - 2*sd(Score),
         Score.q <= mean(Score.q) + 2*sd(Score.q), 
         Score.q >= mean(Score.q) - 2*sd(Score.q),
         year(CreationDate) > 2008)
``` 


## Descriptives

#### What is the distribution of answer scores.

```{r}
answers %>% 
  ggplot(aes(x = Score)) +
  geom_histogram(bins = 20)
```

Scores fall between [-2, 20] with the majority of the scores being between [0,6].

#### What is the distribution of the number of answers for each question. 

```{r}
answers %<>% 
  group_by(ParentId) %>% 
  summarise(n.answers = n()) %>% 
  left_join(answers, ., by = "ParentId")
```

```{r}
answers %>% 
  ggplot(aes(x = n.answers)) +
  geom_histogram(bins = 30)
```

Most questions have fewer than 5 answers and the max is 11. 

#### What proportion of questions have an accepted answer?

```{r}
answers %>% 
  group_by(IsAcceptedAnswer) %>% 
  summarise(prop = n() / nrow(.)) %>% 
  kable(digits = 2)
```

Pretty much an even split between accepted and not accepted answers, with a few more questions without accepted answers. I think this could be both an interesting outcome and an interesting predictor for subsequent analyses.

#### How much time typically passes between a question and a response?

```{r}
answers %<>% 
  mutate(time_diff_sec = as.numeric(CreationDate - CreationDate.q),
         time_diff_min = time_diff_sec / 60, 
         time_diff_hours = time_diff_min / 60, 
         time_diff_days = time_diff_hours / 24)
```

TODO: gather the reduant columns to put this data in tidy format: variables in columns and observations in rows. 

```{r}
answers %<>% gather(key = unit, value = time_diff, ... = time_diff_sec:time_diff_days)
answers %<>% gather(key = unit, value = time_diff, ... = time_diff_sec:time_diff_days)
```




